---
title: 'Homework #5:  Count Regression Models'
author: "Douglas Barley, Ethan Haley, Isabel Magnus, John Mazon, Vinayak Kamath, Arushi"
date: "11/28/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{dcolumn}
---



```{r setup and packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, warning = F)
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
if (!require("Hmisc",character.only = TRUE)) (install.packages("Hmisc",dep=TRUE))
if (!require("ClusterR",character.only = TRUE)) (install.packages("ClusterR",dep=TRUE))
if (!require("cluster",character.only = TRUE)) (install.packages("cluster",dep=TRUE))
if (!require("pscl",character.only = TRUE)) (install.packages("pscl",dep=TRUE))
if (!require("AER",character.only = TRUE)) (install.packages("AER",dep=TRUE))
if (!require("corrgram",character.only = TRUE)) (install.packages("corrgram",dep=TRUE))
if (!require("stargazer",character.only = TRUE)) (install.packages("stargazer",dep=TRUE))
if (!require("ROCR",character.only = TRUE)) (install.packages("ROCR",dep=TRUE))

library(MASS)
library(pscl)
library(AER)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(corrplot)
library(MASS)
library(caret)
library(tidyr)
library(data.table)
require(car)
require(corrgram)
require(ggplot2)
library(ClusterR)
library(cluster)
library(kableExtra)
library(pscl)
library(AER)
library(corrgram)
library(stargazer)
library(ROCR)


```
# OVERVIEW 

In  this  homework  assignment,  you  will  explore,  analyze  and  model  a  data  set  containing  information  on 
approximately 12,000 commercially available wines. The variables are mostly related to the chemical properties of 
the wine being sold. The response variable is the number of sample cases of wine that were purchased by wine 
distribution  companies  after  sampling  a  wine.  These  cases  would  be  used  to  provide  tasting  samples  to 
restaurants  and  wine  stores  around  the  United  States.  The  more  sample  cases  purchased,  the  more  likely  is  a 
wine  to  be  sold  at  a  high  end  restaurant.  A  large  wine  manufacturer  is  studying  the  data  in  order  to  predict  the 
number  of  wine  cases  ordered  based  upon  the  wine  characteristics.  If  the  wine  manufacturer  can  predict  the 
number of cases, then that manufacturer will be able to adjust their wine offering to maximize sales. 
 
Your  objective  is  to  build  a  count  regression  model  to  predict  the  number  of  cases  of  wine  that  will  be  sold 
given certain properties of the wine. HINT: Sometimes, the fact that a variable is missing is actually predictive of 
the target. You can only use the variables given to you (or variables that you derive from the variables provided). 
Below is a short description of the variables of interest in the data set:

-**INDEX**:    Identification Variable (do not use) None

-**TARGET**: Number of Cases Purchased None

-**AcidIndex**: Proprietary method of testing total acidity of wine by using a weighted average

-**Alcohol**: Alcohol Content

-**Chlorides**: Chloride content of wine

-**CitricAcid**: Citric Acid Content

-**Density**: Density of Wine

-**FixedAcidity**: Fixed Acidity of Wine

-**FreeSulfurDioxide**: Sulfur Dioxide content of wine

-**LabelAppeal**:
Marketing Score indicating the appeal of label design for consumers. High numbers 
suggest customers like the label design. Negative numbers suggest customes 
don't like the design.
Many consumers purchase based on the visual appeal of the 
wine label design. Higher numbers suggest better sales.

-**ResidualSugar**: Residual Sugar of wine

-**STARS**: Wine rating by a team of experts. 4 Stars = Excellent, 1 Star = Poor A high number of stars suggests high sales

-**Sulphates**: Sulfate content of wine

-**TotalSulfurDioxide**: Total Sulfur Dioxide of Wine

-**VolatileAcidity**: Volatile Acid content of wine

-**pH**: pH of wine

```{r}
#importing the train an eval data
wine_train_df<- read.csv("https://raw.githubusercontent.com/johnm1990/msds-621/main/wine-training-data.csv")
wine_train_df <- wine_train_df[,2:16]
wine_eval_df<- read.csv("https://raw.githubusercontent.com/johnm1990/msds-621/main/wine-evaluation-data.csv")
wine_eval_df <- wine_eval_df[,2:16]
#per assignment instructions, we don't use first column 'ID', so we remove it, we performed in above manner to keep all rows
```

# DATA EXPLORATION

```{r Section 1}
summary(wine_train_df)
kable(format(sapply(wine_train_df, function(wine_train_df) c( "Stand dev" = round(sd(wine_train_df, na.rm = T),2), 
                         "Mean"= mean(wine_train_df,na.rm=TRUE),
                         "n" = length(wine_train_df),
                         "Median" = median(wine_train_df,na.rm = TRUE),
                         "CoeffofVariation" = sd(wine_train_df)/mean(wine_train_df,na.rm=TRUE),
                         "Minimum" = min(wine_train_df),
                         "Maximum" = max(wine_train_df),
                         "Upper Quantile" = quantile(wine_train_df,1,na.rm = TRUE),
                         "LowerQuartile" = quantile(wine_train_df,0,na.rm = TRUE)
                    )
), scientific = FALSE)
)

ggplot(gather(wine_train_df), aes(value)) + 
    geom_histogram(bins = 10) + 
    facet_wrap(~key, scales = 'free_x')
hist(wine_train_df$TARGET)
table(wine_train_df$TARGET)
#Corr matrix and the scatterplot matrix
##correlation matrix
wine_train_df.rcorr = rcorr(as.matrix(wine_train_df))
wine_train_df.rcorr

wine_train_df.cor = cor(wine_train_df, use = "pairwise.complete.obs")
corrplot(wine_train_df.cor)

histogram(~ TARGET | STARS, data = wine_train_df)
histogram(~ TARGET | LabelAppeal, data = wine_train_df)
histogram(~ AcidIndex | TARGET, data = wine_train_df)
cor_stars_tgt <- cor.test(wine_train_df$STARS, wine_train_df$TARGET)
cor_stars_tgt
cor_lbl_tgr <- cor.test(wine_train_df$LabelAppeal, wine_train_df$TARGET)
cor_lbl_tgr
cor_acid_tgt <- cor.test(wine_train_df$AcidIndex, wine_train_df$TARGET)
cor_acid_tgt

# Compute the analysis of variance, when has more than two groups perform ANOVA
res.aov <- aov(AcidIndex ~ TARGET, data = wine_train_df)
# Summary of the analysis
summary(res.aov)
```

# DATA PREPARATION


```{r section 2}
wine_train_df <- wine_train_df %>% 
            mutate(ResidualSugar= ifelse(is.na(ResidualSugar), 
                                         mean(ResidualSugar, na.rm=TRUE),ResidualSugar),
                   Chlorides= ifelse(is.na(Chlorides), 
                                         mean(Chlorides, na.rm=TRUE),Chlorides),
                   FreeSulfurDioxide= ifelse(is.na(FreeSulfurDioxide), 
                                         mean(FreeSulfurDioxide, na.rm=TRUE),FreeSulfurDioxide),
                   TotalSulfurDioxide= ifelse(is.na(TotalSulfurDioxide), 
                                         mean(TotalSulfurDioxide, na.rm=TRUE),TotalSulfurDioxide),
                   pH= ifelse(is.na(pH), 
                                         mean(pH, na.rm=TRUE),pH),
                   Alcohol= ifelse(is.na(Alcohol), 
                                         mean(Alcohol, na.rm=TRUE),Alcohol),
                   )

wine_eval_df <- wine_eval_df %>% 
            mutate(ResidualSugar= ifelse(is.na(ResidualSugar), 
                                         mean(ResidualSugar, na.rm=TRUE),ResidualSugar),
                   Chlorides= ifelse(is.na(Chlorides), 
                                         mean(Chlorides, na.rm=TRUE),Chlorides),
                   FreeSulfurDioxide= ifelse(is.na(FreeSulfurDioxide), 
                                         mean(FreeSulfurDioxide, na.rm=TRUE),FreeSulfurDioxide),
                   TotalSulfurDioxide= ifelse(is.na(TotalSulfurDioxide), 
                                         mean(TotalSulfurDioxide, na.rm=TRUE),TotalSulfurDioxide),
                   pH= ifelse(is.na(pH), 
                                         mean(pH, na.rm=TRUE),pH),
                   Alcohol= ifelse(is.na(Alcohol), 
                                         mean(Alcohol, na.rm=TRUE),Alcohol),
                   )




#LOG TRANSFORMATION

wine_train_df$FreeSulfurDioxide_log <- log(wine_train_df$FreeSulfurDioxide + 1 - min(wine_train_df$FreeSulfurDioxide))
wine_train_df$TotalSulfurDioxide_log <- log(wine_train_df$TotalSulfurDioxide + 1 - min(wine_train_df$TotalSulfurDioxide))

wine_eval_df$FreeSulfurDioxide_log <- log(wine_eval_df$FreeSulfurDioxide + 1 - min(wine_eval_df$FreeSulfurDioxide))
wine_eval_df$TotalSulfurDioxide_log <- log(wine_eval_df$TotalSulfurDioxide + 1 - min(wine_eval_df$TotalSulfurDioxide))  


                   
# #Flags for N/A's:
    
wine_train_df <- wine_train_df %>% 
            mutate(Sulphates_flag= ifelse(is.na(Sulphates),1,0),
                   STARS_flag= ifelse(is.na(STARS),1,0)
                   )
#flags = will create 1 if NA
histogram(~ TARGET | STARS_flag, data = wine_train_df)
histogram(~ TARGET | Sulphates_flag, data = wine_train_df)

#corrective actions
wine_train_df <- wine_train_df %>% 
            mutate(Sulphates= ifelse(is.na(Sulphates), 
                                         mean(Sulphates, na.rm=TRUE),Sulphates),
                   STARS_merged=ifelse(is.na(STARS),0,STARS))

wine_eval_df <- wine_eval_df %>% 
            mutate(Sulphates= ifelse(is.na(Sulphates), 
                                         mean(Sulphates, na.rm=TRUE),Sulphates),
                   STARS_merged=ifelse(is.na(STARS),0,STARS))

table(wine_train_df$STARS) 
table(wine_train_df$STARS_merged)
##you will see includes no 0 columns
table(wine_train_df$STARS)

#creating clusters for acid index
kmeans.re <- kmeans(wine_train_df$AcidIndex, centers = 5)
table(kmeans.re$cluster)
wine_train_df$AcidIndex_clusters <- kmeans.re$cluster
histogram(~ TARGET | AcidIndex_clusters, data = wine_train_df)

```

# BUILD THE MODELS



```{r}
#multiple reg
model.manual.mr <- lm(TARGET ~ STARS_merged+LabelAppeal+AcidIndex, data = wine_train_df)
summary(model.manual.mr)
#
fullmod_regressiondata <- wine_train_df %>% 
  dplyr::select(TARGET,FixedAcidity,VolatileAcidity,CitricAcid,
    ResidualSugar,Chlorides,Density,pH,Sulphates,Alcohol,LabelAppeal,AcidIndex, 
   FreeSulfurDioxide_log,TotalSulfurDioxide_log, 
    STARS_merged)
#
model.full.mr  <- lm(TARGET ~ . , data = fullmod_regressiondata)
summary(model.full.mr)
model.forward.mr <- model.full.mr %>% stepAIC(direction = "forward", trace = FALSE)
summary(model.forward.mr)
#Getting formula for the model 
formula(model.forward.mr)
model.backward.mr <- model.full.mr %>% stepAIC(direction = "backward", trace = FALSE)
summary(model.backward.mr)
AIC(model.backward.mr)
#Getting formula for the model 
formula(model.backward.mr)
```



```{r}
#manual poisson
model.manual.poisson <- glm(TARGET ~ STARS_merged+LabelAppeal+AcidIndex, data = wine_train_df,family = poisson)
summary(model.manual.poisson)
model.full.poisson  <- glm(TARGET ~ . , data = fullmod_regressiondata,family=poisson)
summary(model.full.poisson)
model.forward.poisson <- model.full.poisson %>% stepAIC(direction = "forward", trace = FALSE)
summary(model.forward.poisson)
#Getting formula for the model 
formula(model.forward.poisson)
model.backward.poisson<-model.full.poisson %>% stepAIC(direction = "backward", trace = FALSE)
summary(model.backward.poisson)
#Getting formula for the model 
formula(model.backward.poisson)
```

Backward consistently provided better results.


```{r}
#negative binomial
model.manual.negbin <- glm.nb(TARGET ~ STARS_merged+LabelAppeal+AcidIndex, data = wine_train_df)
summary(model.manual.negbin)
#Step 1: Create a full model
model.full.negbin  <- glm.nb(TARGET ~ . , data = fullmod_regressiondata)
summary(model.full.negbin )
model.forward.negbin <- model.full.negbin %>% stepAIC(direction = "forward", trace = FALSE)
summary(model.forward.negbin)
#Getting formula for the model 
formula(model.forward.negbin)
model.backward.negbin <-model.full.negbin %>% stepAIC(direction = "backward", trace = FALSE)
summary(model.backward.negbin)
#Getting formula for the model 
formula(model.backward.negbin)
```




# SELECT THE MODELS

```{r, results='asis'}
stargazer(model.full.mr, model.forward.poisson, model.forward.negbin, title="Results", align=TRUE)#, header=FALSE, type='latex')
```



Predictions

```{r}
predict1 <- predict(model.forward.mr, newdata=wine_eval_df, type="response")
summary(predict1)
write.csv(predict1, 'predict1.csv', row.names = FALSE)

```


```{r}
predict2 <- predict(model.forward.poisson, newdata=wine_eval_df, type="response")
summary(predict2)
write.csv(predict2, 'predict2.csv', row.names = FALSE)

```

```{r}
predict3 <- predict(model.forward.negbin, newdata=wine_eval_df, type="response")
summary(predict3)
write.csv(predict3, 'predict3.csv', row.names = FALSE)

```






