---
title: "EvaluationPredictions"
author: "Ethan Haley"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This notebook creates predictions for the evaluation set based on training new models from the entire training set.

```{r}
# load data from Doug's github repo
trainDF = read.csv('https://raw.githubusercontent.com/douglasbarley/DATA621/main/Homework1/moneyball-training-data.csv')
evalDF = read.csv('https://raw.githubusercontent.com/douglasbarley/DATA621/main/Homework1/moneyball-evaluation-data.csv')
```

Change columns

```{r}
# remove meaningless index col
trainDF <- subset(trainDF, select = -c(INDEX))
evalDF <- subset(evalDF, select = -c(INDEX))
# most rows have NA for HBP col, so just make a binary for is.na there, and maybe train on it or maybe not
trainDF$TEAM_BATTING_HBP[!is.na(trainDF$TEAM_BATTING_HBP)] = 0
trainDF$TEAM_BATTING_HBP[is.na(trainDF$TEAM_BATTING_HBP)] = 1

evalDF$TEAM_BATTING_HBP[!is.na(evalDF$TEAM_BATTING_HBP)] = 0
evalDF$TEAM_BATTING_HBP[is.na(evalDF$TEAM_BATTING_HBP)] = 1

singles_hit = trainDF$TEAM_BATTING_H - trainDF$TEAM_BATTING_2B - trainDF$TEAM_BATTING_3B - trainDF$TEAM_BATTING_HR
trainDF$TEAM_BATTING_1B = singles_hit

# We only get Hits and HR for Pitching stats, so can only separate into HR vs all others
hits_allowed = trainDF$TEAM_PITCHING_H - trainDF$TEAM_PITCHING_HR
trainDF$TEAM_PITCHING_1B2B3B = hits_allowed

# Now we can discard the columns we used to make new columns
trainDF = subset(trainDF, select=-c(TEAM_BATTING_H, TEAM_PITCHING_H))

# Repeat for eval data
singles_hit = evalDF$TEAM_BATTING_H - evalDF$TEAM_BATTING_2B - evalDF$TEAM_BATTING_3B - evalDF$TEAM_BATTING_HR
evalDF$TEAM_BATTING_1B = singles_hit

# We only get Hits and HR for Pitching stats, so can only separate into HR vs all others
hits_allowed = evalDF$TEAM_PITCHING_H - evalDF$TEAM_PITCHING_HR
evalDF$TEAM_PITCHING_1B2B3B = hits_allowed

# Now we can discard the columns we used to make new columns
evalDF = subset(evalDF, select=-c(TEAM_BATTING_H, TEAM_PITCHING_H))

dim(trainDF)
dim(evalDF)
```

 Subset into 4 groups each  
 
```{r}
fullTrains = na.omit(trainDF)
fullEvals = na.omit(evalDF)
# .... while the second group has at least one NA in some feature
naTrains = trainDF[rowSums(is.na(trainDF)) > 0, ]
naEvals = evalDF[rowSums(is.na(evalDF)) > 0, ]
naTrains = subset(naTrains, select = -c(TEAM_BATTING_HBP)) # all are NAs
naEvals = subset(naEvals, select = -c(TEAM_BATTING_HBP)) # all are NAs 

#logs
for (column in colnames(fullTrains)){
  fullTrains[paste('log_', column, sep='')] <- log(fullTrains[[column]] + 1)  # +1 for the zeros
}
fullTrains <- subset(fullTrains, select = -c(log_TARGET_WINS, log_TEAM_BATTING_HBP))

for (column in colnames(fullEvals)){
  fullEvals[paste('log_', column, sep='')] <- log(fullEvals[[column]] + 1)  # +1 for the zeros
}
fullEvals <- subset(fullEvals, select = -c(log_TEAM_BATTING_HBP))

for (column in colnames(naTrains)){
  naTrains[paste('log_', column, sep='')] <- log(naTrains[[column]] + 1)  # +1 for the zeros
}
naTrains <- subset(naTrains, select = -c(log_TARGET_WINS))

for (column in colnames(naEvals)){
  naEvals[paste('log_', column, sep='')] <- log(naEvals[[column]] + 1)  # +1 for the zeros
}
```

```{r}
# Make NA binary indicators for a few columns
naTrains$TEAM_BASERUN_SB[!is.na(naTrains$TEAM_BASERUN_SB)] = 0
naTrains$TEAM_BASERUN_SB[is.na(naTrains$TEAM_BASERUN_SB)] = 1

naTrains$TEAM_BASERUN_CS[!is.na(naTrains$TEAM_BASERUN_CS)] = 0
naTrains$TEAM_BASERUN_CS[is.na(naTrains$TEAM_BASERUN_CS)] = 1

naTrains$TEAM_FIELDING_DP[!is.na(naTrains$TEAM_FIELDING_DP)] = 0
naTrains$TEAM_FIELDING_DP[is.na(naTrains$TEAM_FIELDING_DP)] = 1

```

```{r}
# Make NA binary indicators for a few columns
naEvals$TEAM_BASERUN_SB[!is.na(naEvals$TEAM_BASERUN_SB)] = 0
naEvals$TEAM_BASERUN_SB[is.na(naEvals$TEAM_BASERUN_SB)] = 1

naEvals$TEAM_BASERUN_CS[!is.na(naEvals$TEAM_BASERUN_CS)] = 0
naEvals$TEAM_BASERUN_CS[is.na(naEvals$TEAM_BASERUN_CS)] = 1

naEvals$TEAM_FIELDING_DP[!is.na(naEvals$TEAM_FIELDING_DP)] = 0
naEvals$TEAM_FIELDING_DP[is.na(naEvals$TEAM_FIELDING_DP)] = 1

```

Now split the data with NA's

```{r}
naSO = naTrains[is.na(naTrains$TEAM_BATTING_SO), ] # SO is NA
noSO = naTrains[(!is.na(naTrains$TEAM_BATTING_SO)) & (naTrains$TEAM_BATTING_SO == 0), ] # SO is 0
naTrains = naTrains[(!is.na(naTrains$TEAM_BATTING_SO)) & (naTrains$TEAM_BATTING_SO > 0), ] # all the rest
```

```{r}
naSOeval = naEvals[is.na(naEvals$TEAM_BATTING_SO), ] # SO is NA
noSOeval = naEvals[(!is.na(naEvals$TEAM_BATTING_SO)) & (naEvals$TEAM_BATTING_SO == 0), ] # SO is 0
naEvals = naEvals[(!is.na(naEvals$TEAM_BATTING_SO)) & (naEvals$TEAM_BATTING_SO > 0), ] # all the rest
```

For full model add HR/SO ratio 

```{r}
# Create HR-to-SO ratio columns
fullTrains$HR2SOfor = fullTrains$TEAM_BATTING_HR / fullTrains$TEAM_BATTING_SO
fullTrains$HR2SOvs  = fullTrains$TEAM_PITCHING_HR / fullTrains$TEAM_PITCHING_SO
```

```{r}
# Create HR-to-SO ratio columns
fullEvals$HR2SOfor = fullEvals$TEAM_BATTING_HR / fullEvals$TEAM_BATTING_SO
fullEvals$HR2SOvs  = fullEvals$TEAM_PITCHING_HR / fullEvals$TEAM_PITCHING_SO
```



```{r}
model2 = lm(TARGET_WINS ~ . - TEAM_BATTING_SO - TEAM_BATTING_1B - log_TEAM_BASERUN_SB - 
              log_TEAM_FIELDING_E - TEAM_PITCHING_BB - TEAM_PITCHING_HR - 
              log_TEAM_BATTING_1B - log_TEAM_PITCHING_1B2B3B - TEAM_FIELDING_DP -
              log_TEAM_BATTING_HR, fullTrains)
summary(model2)
```

 ==============
 Full Eval Preds 
 ==============
 
```{r}
m2preds = predict(model2, fullEvals)
m2preds
```
 
 Next up is the remaining data whose SO values are positive integers  
 but some other predictor is NA.  [naTrains and naEvals]  
 

```{r}
#remove constructed log cols with NA's
naTrains = subset(naTrains, select = -c(log_TEAM_BASERUN_SB, log_TEAM_BASERUN_CS, log_TEAM_FIELDING_DP))
summary(naTrains)
```


- TEAM_BATTING_HR - TEAM_PITCHING_HR - TEAM_BATTING_BB - 
              TEAM_FIELDING_DP - TEAM_BATTING_SO - TEAM_PITCHING_SO

```{r}
model3 = lm(TARGET_WINS ~ . - TEAM_FIELDING_DP - log_TEAM_BATTING_3B - TEAM_PITCHING_1B2B3B - log_TEAM_PITCHING_BB - log_TEAM_BATTING_HR - log_TEAM_BATTING_BB - TEAM_BATTING_BB - TEAM_BATTING_HR - TEAM_BATTING_2B - TEAM_BATTING_1B - log_TEAM_PITCHING_HR - TEAM_BATTING_3B - 
              TEAM_PITCHING_HR, naTrains)
summary(model3)
```

 ==============
 Pos SO NA other Preds 
 ==============

```{r}
m3preds = predict(model3, naEvals)
m3preds
```

Now naSO and naSOeval  (NA SO values)

```{r}
naSO = subset(naSO, select = -c(TEAM_BATTING_SO, TEAM_PITCHING_SO, log_TEAM_BATTING_SO, 
                                log_TEAM_BASERUN_CS, log_TEAM_PITCHING_SO))
naSO = subset(naSO, select = -c(TEAM_FIELDING_DP, TEAM_BASERUN_SB, TEAM_BASERUN_CS))
summary(naSO)
```

```{r}
model4 = lm(TARGET_WINS ~ ., naSO)
summary(model4)
```

 ==============
 NA SO Preds 
 ==============
 
```{r}
m4preds = predict(model4, naSOeval)
m4preds
```
 
Last, zero SO data  

```{r}
noSO$invHitsAllowed = 1 / noSO$TEAM_PITCHING_1B2B3B
noSOeval$invHitsAllowed = 1 / noSOeval$TEAM_PITCHING_1B2B3B
noSO = subset(noSO, select = -c(TEAM_BATTING_SO, TEAM_PITCHING_SO, TEAM_FIELDING_DP, 
                                log_TEAM_BATTING_SO, log_TEAM_BASERUN_SB, log_TEAM_BASERUN_CS, 
                                log_TEAM_PITCHING_SO, log_TEAM_FIELDING_DP))
summary(noSO)
```

```{r}
model5 = lm(TARGET_WINS ~ . - log_TEAM_BATTING_2B - TEAM_BASERUN_CS - log_TEAM_BATTING_3B - 
              log_TEAM_PITCHING_HR - TEAM_BATTING_3B - TEAM_BATTING_2B - log_TEAM_BATTING_BB - 
              TEAM_BATTING_1B - TEAM_PITCHING_1B2B3B - log_TEAM_PITCHING_1B2B3B, noSO)
summary(model5)
```

 ==============
 Zero SO Preds 
 ==============

```{r}
m5preds = predict(model5, noSOeval)
m5preds
```

Clip to 0 wins predicted
 
```{r}
m5preds = pmax(m5preds, 0)
m5preds
```

Patch the 4 predictions back in the necessary order:  

```{r}
evals = data.frame(ind=as.integer(c(names(m2preds), names(m3preds), names(m4preds), names(m5preds))), 
                   pred=c(m2preds,m3preds,m4preds,m5preds))
```

```{r}
evals
```

```{r}
evals = evals[order(evals$ind), ]
evals
```
```{r}
evals = subset(evals, select = c(pred))
evals
```
```{r}
write.csv(evals, "evalPredsMoneyball.csv")
```



