---
title: "Homework #2 Classification Metrics"
author: "Douglas Barley, Ethan Haley, Isabel Magnus, John Mazon, Vinayak Kamath"
date: "9/28/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

 DATA 621 – Business Analytics and Data Mining
 
Homework #2 Assignment Requirements  

 Overview  
 
In this homework assignment, you will work through various classification metrics. You will be asked to create functions in R to carry out the various calculations. You will also investigate some functions in packages that will let you obtain the equivalent results. Finally, you will create graphical output that also can be used to evaluate the output of classification models, such as binary logistic regression.  

Supplemental Material
 - Applied Predictive Modeling, Ch. 11 (provided as a PDF file).
 - Web tutorials: http://www.saedsayad.com/model_evaluation_c.htm
Deliverables (100 Points)
 - Upon following the instructions below, use your created R functions and the other packages to generate the classification metrics for the provided data set. A write-up of your solutions submitted in PDF format.   

Instructions
Complete each of the following steps as instructed:  

1. Download the classification output data set (attached in Blackboard to the assignment).   

```{r}
d.f = read.csv('https://raw.githubusercontent.com/ebhtra/msds-621/main/HW2/classification-output-data.csv')
head(d.f)
colnames(d.f)
```



2. The data set has three key columns we will use:  

 - **class**: the actual class for the observation
 - **scored.class**: the predicted class for the observation (based on a threshold of 0.5)
 - **scored.probability**: the predicted probability of success for the observation
Use the **table()** function to get the raw confusion matrix for this scored dataset. Make sure you understand the output. In particular, do the rows represent the actual or predicted class? The columns?   

```{r}
data1 <- d.f %>% 
  select(class, scored.class, scored.probability)
head(data1)

```

A confusion matrix shows the number of correct and incorrect predictions made by the classification model compared to the actual outcomes (target value) in the data. The matrix is NxN, where N is the number of target values (classes). Performance of such models is commonly evaluated using the data in the matrix. The following table displays a 2x2 confusion matrix for two classes (Positive and Negative).

```{r}
cm_df <- dplyr::select(data1, scored.class, class)
table(cm_df)
```


#### Before we calculate all the metrics, we can build the pieces  



```{r}
# Calculate all the pieces we need
FP = function(d.f, predicted, actual) sum(d.f[predicted] - d.f[actual] == 1)
TP = function(d.f, predicted, actual) sum(d.f[predicted] * d.f[actual] == 1)
FN = function(d.f, predicted, actual) sum(d.f[predicted] - d.f[actual] == -1)
TN = function(d.f, predicted, actual) sum(d.f[predicted] + d.f[actual] == 0) 
```

```{r}
# test above funcs vs the confusion matrix 
FP(d.f, 'scored.class', 'class')
TP(d.f, 'scored.class', 'class')
FN(d.f, 'scored.class', 'class')
TN(d.f, 'scored.class', 'class')
```



3. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the accuracy of the predictions.  

Accuracy = $\frac{TP+TN}{TP+FP+TN+FN}$

```{r}

```


4. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the classification error rate of the predictions. 

Classification Error Rate = $\frac{FP+FN}{TP+FP+TN+FN}$


```{r}

```


Verify that you get an accuracy and an error rate that sums to one.  

```{r}

```



5. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the precision of the predictions.  

Precision = $\frac{TP}{TP+FP}$

```{r}

```



6. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the sensitivity of the predictions. Sensitivity is also known as recall.  

Sensitivity = $\frac{TP}{TP+FN}$  

```{r}

```


7. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the specificity of the predictions.  

Specificity = $\frac{TN}{FP+TN}$  

```{r}

```



8. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the F1 score of the predictions. 

$F1~ Score = \frac{2 \times Precision \times Sensitivity}{Precision + Sensitivity}$  

```{r}

```



9. Before we move on, let’s consider a question that was asked: What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0 < $\alpha$ < 1 and 0 < $\beta$ < 1 then $\alpha \cdot \beta$ < $\alpha$)  


$F1~ Score = \frac{2 \times Precision \times Sensitivity}{Precision + Sensitivity}$, where Precision = $\frac{TP}{TP+FP}$ and Sensitivity = $\frac{TP}{TP+FN}$  

All of TP, FP, and FN are non-negative integers by definition, so we don't have to consider negative values here.

If TP equals 0, then Precision and Sensitivity are either 0 or undefined, and in either case, the F1 Score is undefined. 

All remaining possibilities involve TP being a positive integer.  

The $F1~ Score = \frac{2 \times \frac{TP}{TP+FP} \times \frac{TP}{TP+FN}}{\frac{TP}{TP+FP} + \frac{TP}{TP+FN}} = \frac{2\cdot TP}{2\cdot TP+FN+FP}$  

If both FN and FP are equal to 0, then we get the upper bound on F1, which is 1.  
The lower bound is if TP is equal to 1, in which case we get $\frac{2}{2+FN+FP}$, which approaches 0 as FN and/or FP approach infinity.  Thus, $0\lt F1 \le 1$ if F1 is defined.  Note that a tighter lower bound is provided by the number of predictions, $n$:  $\frac{2}{1+n}$, since $FN+FP \le n-1$.  




10. Write a function that generates an ROC curve from a data set with a true classification column (class in our example) and a probability column (scored.probability in our example). Your function should return a list that includes the plot of the ROC curve and a vector that contains the calculated area under the curve (AUC). Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals.   

```{r}

```



11. Use your created R functions and the provided classification output data set to produce all of the classification metrics discussed above.   

```{r}

```



12. Investigate the caret package. In particular, consider the functions confusionMatrix, sensitivity, and specificity. Apply the functions to the data set. How do the results compare with your own functions?  

```{r}

```



13. Investigate the pROC package. Use it to generate an ROC curve for the data set. How do the results compare with your own functions?  

   
```{r}

```






   