---
title: "Homework #3 Binary  logistic  regression  models "
author: "Douglas Barley, Ethan Haley, Isabel Magnus, John Mazon, Vinayak Kamath, Arushi Arora"
date: "11/1/2021"
output:
  html_document: 
    toc: true
    toc-title: "Homework #3 - Binary logistic regression models"
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: darkly
    highlight: pygments
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
library(ggplot2)
library(knitr)
library(xtable)
library(dplyr)
library(stringr)
library(tidyverse)
library(dplyr)
library(ROCR)
library(Hmisc)
library(corrplot)
library(MASS)
library(caret)
library(data.table)
require(data.table)
require(car)
require(corrgram)
require(ggplot2)
```

 DATA 621 – Business Analytics and Data Mining

# Overview 

Homework #3 Assignment Requirements  
In this homework assignment, you will explore, analyze and model a data set containing information on crime 
for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime 
rate is above the median crime rate (1) or not (0). 
 
Your  objective  is  to  build  a  binary  logistic  regression  model  on  the  training  data  set  to  predict  whether  the 
neighborhood  will  be  at  risk  for  high  crime  levels.  You  will  provide  classifications  and  probabilities  for  the 
evaluation data  set using  your  binary  logistic  regression  model. You  can only  use  the  variables  given  to  you (or 
variables that you derive from the variables provided). Below is a short description of the variables of interest in 
the data set


 
```{r libraries & reading data}
crime_eval_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-evaluation-data_modified.csv")
crime_train_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-training-data_modified.csv")
```



## 1.  Data Exploration



```{r}
head(crime_train_df, 1)
```




```{r}
###summary statistics
summary(crime_train_df)
sapply(crime_train_df, sd, na.rm=TRUE)
sapply(crime_train_df, hist, na.rm=TRUE)

##correlation matrix
crime_train_df.rcorr = rcorr(as.matrix(crime_train_df))
crime_train_df.rcorr

crime_train_df.cor = cor(crime_train_df)
corrplot(crime_train_df.cor)
#correlation_matrix = crime_train_df.corr()


#correlation plot
crime_train_df.cor = cor(crime_train_df)
corrplot(crime_train_df.cor)

cor_distarget <- cor.test(crime_train_df$dis, crime_train_df$target)
cor_distarget

cor_noxtarget <- cor.test(crime_train_df$nox, crime_train_df$target)
cor_noxtarget

cor_taxrad<- cor.test(crime_train_df$tax, crime_train_df$rad)
cor_taxrad

#ttests to look at difference in means between target = 0 and target = 1, maybe to help create buckets
Ttest_targetpt<- t.test(crime_train_df$ptratio ~ crime_train_df$target)
Ttest_targetpt

Ttest_targetnox<- t.test(crime_train_df$nox ~ crime_train_df$target)
Ttest_targetnox

Ttest_targetpt<- t.test(crime_train_df$ptratio ~ crime_train_df$target)
Ttest_targetpt

Ttest_targetlstat<- t.test(crime_train_df$lstat ~ crime_train_df$target)
Ttest_targetlstat


```

## 2.  Data Preparation

```{r}
crime_train_df$target <-  as.factor(crime_train_df$target)
boxplot(crime_train_df$zn ~ crime_train_df$target)
hist(crime_train_df$zn)
table(crime_train_df$zn[crime_train_df$zn<10])
hist(crime_train_df$zn[crime_train_df$zn>0])
table(crime_train_df$target[crime_train_df$zn > 22])
crime_train_df['zn_hi'] = crime_train_df$zn > 22
boxplot(crime_train_df$indus ~ crime_train_df$target)
hist(crime_train_df$indus)
table(crime_train_df$indus[crime_train_df$indus > 18])
table(crime_train_df$target[(crime_train_df$indus>18) & (crime_train_df$indus<20)])
table(crime_train_df$target[(crime_train_df$indus>20)])


#Make a dummy indicator for 18<indus<20  

crime_train_df['indus19'] = crime_train_df$indus > 18 & crime_train_df$indus < 20
 

boxplot(crime_train_df$nox ~ crime_train_df$target)
boxplot(crime_train_df$rm ~ crime_train_df$target)

boxplot(crime_train_df$age ~ crime_train_df$target)
hist(crime_train_df$age[crime_train_df$target==0])

#That looks normal for the lower crime areas.  How about the higher ones?  

hist(crime_train_df$age[crime_train_df$target==1])
crime_train_df['sq_age'] = crime_train_df$age ** 2
boxplot(crime_train_df$dis ~ crime_train_df$target)
sd(crime_train_df$dis[crime_train_df$target==1])
sd(crime_train_df$dis[crime_train_df$target==0])
hist(log(crime_train_df$dis))

hist(log(crime_train_df$dis[crime_train_df$target==0]))
hist(log(crime_train_df$dis[crime_train_df$target==1]))

#Taking the log normalizes the predictor conditioned on the response, so log_dis should be useful.

crime_train_df['log_dis'] = log(crime_train_df$dis)
boxplot(crime_train_df$rad ~ crime_train_df$target)

sd(crime_train_df$rad[crime_train_df$target==1])
sd(crime_train_df$rad[crime_train_df$target==0])

hist(crime_train_df$rad)

table(crime_train_df$target[crime_train_df$rad>15])

mean(crime_train_df$indus19[crime_train_df$rad>8])

 
#Yes, that's just redundancy.  What about at the lower end of the range? 
 
mean(crime_train_df$indus19[crime_train_df$rad<5])

crime_train_df['rad5to8'] = 5 < crime_train_df$rad & crime_train_df$rad < 8

boxplot(crime_train_df$tax ~ crime_train_df$target)

sd(crime_train_df$tax[crime_train_df$target==1])
sd(crime_train_df$tax[crime_train_df$target==0])


#Very different variances

hist(crime_train_df$tax)
crime_train_df['log_tax'] = log(crime_train_df$tax)
table(crime_train_df$tax[crime_train_df$tax>600])

#That's just weird, why 666?  

table(crime_train_df$target[crime_train_df$tax == 666])


#That's the same table as rad > 15.  More redundancy probably.  

sum(crime_train_df$tax==666 & crime_train_df$rad>15)


#What about the tax == 711:


table(crime_train_df$target[crime_train_df$tax==711])

#Will indus19 take care of those 5, so that we can ignore tax > 600 ?  

sum(crime_train_df$tax==711 & crime_train_df$indus19)

table(crime_train_df$target[crime_train_df$tax==711 & !crime_train_df$indus19])

crime_train_df['tax_666'] = crime_train_df$tax==666


boxplot(crime_train_df$ptratio ~ crime_train_df$target)
hist(crime_train_df$ptratio[crime_train_df$target==1])
table(crime_train_df$ptratio[crime_train_df$ptratio>19])
table(crime_train_df$target[crime_train_df$ptratio==20.2])

crime_train_df['pt_peak'] = crime_train_df$ptratio == 20.2
crime_train_df['log_ptrat'] = log(crime_train_df$ptratio)


boxplot(crime_train_df$lstat ~ crime_train_df$target)

hist(crime_train_df$lstat)


crime_train_df['log_lstat'] = log(crime_train_df$lstat)

boxplot(crime_train_df$medv ~ crime_train_df$target)
```




## 3. Build the Models

```{r}

set.seed(123)
split <- caret::createDataPartition(crime_train_df$target, p=0.90, list=FALSE)
train <- crime_train_df[split, ]
validation <- crime_train_df[ -split, ]
#Step 1: Create a full model
model.full  <- glm(target ~ . , data = train, family = 'binomial')
summary(model.full)
```


```{r}
#Step 2: Create a backward model using the full model 
model.backward <- model.full %>% stepAIC(direction = "backward", trace = FALSE)
summary(model.backward)
```


```{r}
#Getting formula for the model 
formula(model.backward)
```




## 4. Select Models

```{r}
# generating the predictors
model.backward.pred =predict(model.backward, newdata = validation)
model.backward.pred[model.backward.pred >= 0.5] <- 1
model.backward.pred[model.backward.pred < 0.5] <- 0
model.backward.pred = as.factor(model.backward.pred)


#  generating the confusion matrix
model.backward.confusion.matrix <- confusionMatrix(model.backward.pred, validation$target, mode = "everything")
model.backward.confusion.matrix

```


```{r}
par(mfrow=c(1,1))
hvalues <- influence(model.backward)$hat
stanresDeviance <- residuals(model.backward)/sqrt(1-hvalues)
plot(hvalues,stanresDeviance,ylab="Standardized Deviance Residuals",xlab="Leverage Values",ylim=c(-3,3),xlim=c(-0.05,0.7))
abline(v=2*7/length(train),lty=2)
identify(hvalues,stanresDeviance,cex=0.75)
```
Check the full dataset fit with the backward-fit last model.

```{r}
fullFit = glm(formula = target ~ zn + nox + rm + age + dis + rad + tax + 
    ptratio + lstat + medv + zn_hi + indus19 + log_dis + rad5to8 + 
    log_tax + log_ptrat + log_lstat, family = "binomial", data = crime_train_df)
summary(fullFit)
```



```{r}


p <- predict(model.backward, type = "response")
roc_pred <- prediction(predictions = p,labels=model.backward$y)

auc.tmp <- performance(roc_pred,"auc"); auc <- as.numeric(auc.tmp@y.values)
auc

#plotting roc
roc_perf <- performance(roc_pred , "tpr" , "fpr")
plot(roc_perf,
     colorize = TRUE,
     print.cutoffs.at= seq(0,1,0.05),
     text.adj=c(-0.2,1.7))


```

