---
title: "Homework #3 Binary  logistic  regression  models "
author: "Douglas Barley, Ethan Haley, Isabel Magnus, John Mazon, Vinayak Kamath"
date: "10/11/2021"
output:
  html_document: 
    toc: true
    toc-title: "Homework #3 - Binary logistic regression models"
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: darkly
    highlight: pygments
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))

library(ggplot2)
library(knitr)
library(xtable)
library(dplyr)
library(stringr)
library(tidyverse)
library(dplyr)
```

 DATA 621 – Business Analytics and Data Mining
 
Homework #3 Assignment Requirements  
In this homework assignment, you will explore, analyze and model a data set containing information on crime 
for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime 
rate is above the median crime rate (1) or not (0). 
 
Your  objective  is  to  build  a  binary  logistic  regression  model  on  the  training  data  set  to  predict  whether  the 
neighborhood  will  be  at  risk  for  high  crime  levels.  You  will  provide  classifications  and  probabilities  for  the 
evaluation data  set using  your  binary  logistic  regression  model. You  can only  use  the  variables  given  to  you (or 
variables that you derive from the variables provided). Below is a short description of the variables of interest in 
the data set

# Overview  
 
```{r libraries & reading data}

crime_eval_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-evaluation-data_modified.csv")
crime_train_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-training-data_modified.csv")

```

```{r section1}
# create summaries for every variable showing basic summaries + NAs

### add the standard deviations to our summaries
summary(crime_train_df)
# Scatterplots between the independent variables and the # wins


##matrix of scatterplots

###Change the scale
# Simple Bar Plot, adjust the scale of the bar plot
counts <- table(crime_train_df$target)
barplot(counts, main="Crime Distribution", 
   xlab="Number of Neighborhoods")

#scatterplots for the target and predictors
pairs(~target + dis +  lstat + ptratio ,
      pch = 19, data = crime_train_df)

```

```{r section 2}

crime_train_df$log_tax <- log(crime_train_df$tax)
summary(crime_train_df)

crime_train_df$chas <-  as.factor(crime_train_df$chas)
crime_train_df$target <-  as.factor(crime_train_df$target)

# creating buckets based on the number of wins to explore the data
crime_train_df$statbuk <- as.numeric(cut_number(crime_train_df$lstat,5))
table(crime_train_df$statbuk)

#check if high SDs, then transform

```
