---
title: "Homework #3 Binary  logistic  regression  models "
author: "Arushi, Douglas, Ethan, Isabel, John, Vinayak"
date: "10/11/2021"
output:
  html_document: 
    toc: true
    toc-title: "Homework #3 - Binary logistic regression models"
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: darkly
    highlight: pygments
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F)
if (!require("ggplot2",character.only = TRUE)) (install.packages("ggplot2",dep=TRUE))
if (!require("knitr",character.only = TRUE)) (install.packages("knitr",dep=TRUE))
if (!require("xtable",character.only = TRUE)) (install.packages("xtable",dep=TRUE))
if (!require("dplyr",character.only = TRUE)) (install.packages("dplyr",dep=TRUE))
if (!require("stringr",character.only = TRUE)) (install.packages("stringr",dep=TRUE))
library(ggplot2)
library(knitr)
library(xtable)
library(dplyr)
library(stringr)
library(tidyverse)
library(dplyr)
```

 DATA 621 – Business Analytics and Data Mining
 
Homework #3 Assignment Requirements  
In this homework assignment, you will explore, analyze and model a data set containing information on crime 
for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime 
rate is above the median crime rate (1) or not (0). 
 
Your  objective  is  to  build  a  binary  logistic  regression  model  on  the  training  data  set  to  predict  whether  the 
neighborhood  will  be  at  risk  for  high  crime  levels.  You  will  provide  classifications  and  probabilities  for  the 
evaluation data  set using  your  binary  logistic  regression  model. You  can only  use  the  variables  given  to  you (or 
variables that you derive from the variables provided). Below is a short description of the variables of interest in 
the data set



# Overview  
 
```{r libraries & reading data}
crime_eval_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-evaluation-data_modified.csv")
crime_train_df <- read.csv("https://raw.githubusercontent.com/johnm1990/DATA621/main/hw3/crime-training-data_modified.csv")
```


```{r}
head(crime_train_df, 1)
```
```{r}
crime_train_df$target <-  as.factor(crime_train_df$target)
```


Go through the predictors in order.

```{r}
boxplot(crime_train_df$zn ~ crime_train_df$target)
```

```{r}
hist(crime_train_df$zn)
```
How many different values are there in that below 10 range?

```{r}
table(crime_train_df$zn[crime_train_df$zn<10])
```

What does the rest of the histogram look like?

```{r}
hist(crime_train_df$zn[crime_train_df$zn>0])
```
Is there a cutoff where every value corresponds to the same target?

```{r}
table(crime_train_df$target[crime_train_df$zn > 22])
```

```{r}
crime_train_df['zn_hi'] = crime_train_df$zn > 22
```



```{r}
boxplot(crime_train_df$indus ~ crime_train_df$target)
```

```{r}
hist(crime_train_df$indus)
```
Is that just one value dominating the histogram?

```{r}
table(crime_train_df$indus[crime_train_df$indus > 18])
```

Why are there so many exact values in the 18-20 area?  Apparently those directly determine the target:  

```{r}
table(crime_train_df$target[(crime_train_df$indus>18) & (crime_train_df$indus<20)])
```

```{r}
table(crime_train_df$target[(crime_train_df$indus>20)])
```

Make a dummy indicator for 18<indus<20  

```{r}
crime_train_df['indus19'] = crime_train_df$indus > 18 & crime_train_df$indus < 20
```


```{r}
boxplot(crime_train_df$nox ~ crime_train_df$target)
```
Leave that for now at least.

```{r}
boxplot(crime_train_df$rm ~ crime_train_df$target)
```

Looks normal, for this dataset.  

```{r}
boxplot(crime_train_df$age ~ crime_train_df$target)
```

```{r}
hist(crime_train_df$age[crime_train_df$target==0])
```

That looks normal for the lower crime areas.  How about the higher ones?  

```{r}
hist(crime_train_df$age[crime_train_df$target==1])
```


```{r}
crime_train_df['sq_age'] = crime_train_df$age ** 2
```


```{r}
boxplot(crime_train_df$dis ~ crime_train_df$target)
```

```{r}
sd(crime_train_df$dis[crime_train_df$target==1])
sd(crime_train_df$dis[crime_train_df$target==0])
```

```{r}
hist(log(crime_train_df$dis))
```

```{r}
hist(log(crime_train_df$dis[crime_train_df$target==0]))
```

```{r}
hist(log(crime_train_df$dis[crime_train_df$target==1]))
```

Taking the log normalizes the predictor conditioned on the response, so log_dis should be useful.

```{r}
crime_train_df['log_dis'] = log(crime_train_df$dis)
```


```{r}
boxplot(crime_train_df$rad ~ crime_train_df$target)
```
```{r}
sd(crime_train_df$rad[crime_train_df$target==1])
sd(crime_train_df$rad[crime_train_df$target==0])
```

```{r}
hist(crime_train_df$rad)
```

```{r}
table(crime_train_df$target[crime_train_df$rad>15])
```

Does this maybe just overlap with indus19 though?

```{r}
mean(crime_train_df$indus19[crime_train_df$rad>8])
```
 
Yes, that's just redundancy.  What about at the lower end of the range? 
 
```{r}
mean(crime_train_df$indus19[crime_train_df$rad<5])
```


So the only values of rad that aren't redundant info, after adding indus 19, are 5<rad<8

```{r}
crime_train_df['rad5to8'] = 5 < crime_train_df$rad & crime_train_df$rad < 8
```


```{r}
boxplot(crime_train_df$tax ~ crime_train_df$target)
```

```{r}
sd(crime_train_df$tax[crime_train_df$target==1])
sd(crime_train_df$tax[crime_train_df$target==0])
```

Very different variances

```{r}
hist(crime_train_df$tax)
```

```{r}
crime_train_df['log_tax'] = log(crime_train_df$tax)
```


```{r}
table(crime_train_df$tax[crime_train_df$tax>600])
```

That's just weird, why 666?  

```{r}
table(crime_train_df$target[crime_train_df$tax == 666])
```

That's the same table as rad > 15.  More redundancy probably.  

```{r}
sum(crime_train_df$tax==666 & crime_train_df$rad>15)
```

What about the tax == 711:

```{r}
table(crime_train_df$target[crime_train_df$tax==711])
```

Will indus19 take care of those 5, so that we can ignore tax > 600 ?  

```{r}
sum(crime_train_df$tax==711 & crime_train_df$indus19)
```

```{r}
table(crime_train_df$target[crime_train_df$tax==711 & !crime_train_df$indus19])
```

```{r}
crime_train_df['tax_666'] = crime_train_df$tax==666
```



```{r}
boxplot(crime_train_df$ptratio ~ crime_train_df$target)
```
```{r}
hist(crime_train_df$ptratio[crime_train_df$target==1])
```
```{r}
table(crime_train_df$ptratio[crime_train_df$ptratio>19])
```

```{r}
table(crime_train_df$target[crime_train_df$ptratio==20.2])
```

```{r}
crime_train_df['pt_peak'] = crime_train_df$ptratio == 20.2
crime_train_df['log_ptrat'] = log(crime_train_df$ptratio)
```


```{r}
boxplot(crime_train_df$lstat ~ crime_train_df$target)
```

```{r}
hist(crime_train_df$lstat)
```

```{r}
crime_train_df['log_lstat'] = log(crime_train_df$lstat)
```



```{r}
boxplot(crime_train_df$medv ~ crime_train_df$target)
```

```{r}
names(crime_train_df)
```

```{r}
library(MASS)
library(caret)

set.seed(123)
split <- caret::createDataPartition(crime_train_df$target, p=0.90, list=FALSE)

train <- crime_train_df[split, ]
validation <- crime_train_df[ -split, ]

#Step 1: Create a full model
model.full  <- glm(target ~ . , data = train, family = 'binomial')
summary(model.full)
```


```{r}
#Step 2: Create a backward model using the full model 
model.backward <- model.full %>% stepAIC(direction = "backward", trace = FALSE)
summary(model.backward)
```


```{r}
#Getting formula for the model 
formula(model.backward)
```


```{r}
# generating the predictors
model.backward.pred =predict(model.backward, newdata = validation)
model.backward.pred[model.backward.pred >= 0.5] <- 1
model.backward.pred[model.backward.pred < 0.5] <- 0
model.backward.pred = as.factor(model.backward.pred)

#  generating the confusion matrix
model.backward.confusion.matrix <- confusionMatrix(model.backward.pred, validation$target, mode = "everything")

model.backward.confusion.matrix

```

```{r}
#library(alr3)
#mmps(model.backward)
```

```{r}
par(mfrow=c(1,1))
hvalues <- influence(model.backward)$hat
stanresDeviance <- residuals(model.backward)/sqrt(1-hvalues)
plot(hvalues,stanresDeviance,ylab="Standardized Deviance Residuals",xlab="Leverage Values",ylim=c(-3,3),xlim=c(-0.05,0.7))
abline(v=2*7/length(train),lty=2)
identify(hvalues,stanresDeviance,cex=0.75)
```
Check the full dataset fit with the backward-fit last model.

```{r}
fullFit = glm(formula = target ~ zn + nox + rm + age + dis + rad + tax + 
    ptratio + lstat + medv + zn_hi + indus19 + log_dis + rad5to8 + 
    log_tax + log_ptrat + log_lstat, family = "binomial", data = crime_train_df)
summary(fullFit)
```
