---
title: "checkoutWHO"
author: "EH"
date: "11/9/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
who = read_csv('../../DATA621/Life Expectancy Data.csv.xls')
```

```{r}
head(who)
```

```{r}
summary(who)
```
Variables in the data, with descriptions

- **Country**  
- **Year**
- **Status**:  Developed or developing
- **Adult Mortality**:  Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)  
- **infant deaths**:  Number of Infant Deaths per 1000 population  
- **Alcohol**:  Alcohol, recorded per capita (15+) consumption (in litres of pure alcohol)  
- **percentage expenditure**:  Expenditure on health as a percentage of Gross Domestic Product per capita (%)  
- **Hepatitis B**:  Hepatitis B (HepB) immunization coverage among 1-year-olds (%)  
- **Measles**:  Measles - number of reported cases per 1000 population  
- **BMI**:  Average Body Mass Index of entire population  
- **under-five deaths**:  Number of under-five deaths per 1000 population  
- **Polio**:  Polio (Pol3) immunization coverage among 1-year-olds (%)  
- **Total expenditure**:  General government expenditure on health as a percentage of total government expenditure (%)  
- **Diphtheria**:  Diphtheria tetanus toxoid and pertussis (DTP3) immunization coverage among 1-year-olds (%)  
- **HIV/AIDS**:  Deaths per 1000 live births HIV/AIDS (0-4 years)  
- **GDP**:  Gross Domestic Product per capita (in USD)  
- **Population**:  Population of the country  
- **thinness 1-19 years**:  Prevalence of thinness among children and adolescents for Age 10 to 19 (% )  
- **thinness 5-9 years**:  Prevalence of thinness among children for Age 5 to 9 (%)  
- **Income composition of resources**:  Human Development Index in terms of income composition of resources (index ranging from 0 to 1)  
- **Schooling**:  Number of years of Schooling  


